---
title: "ASSIGNMENT_2_FML - 811287455"
author: "Sudarshan rayapati"
date: "2023-09-30"
output:
  html_document: default
  pdf_document: default
---




## Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.

Partition the data into training (60%) and validation (40%) sets

***
## installing the pacakges "class","caret","e1071"

## calling the libraries "class","caret","e1071"



```{r}
library(class)
library(caret)
library(e1071)

```


## Reading the bank csv file  

```{r}
x<-read.csv("/Users/sudarshan/Desktop/FML/dataset/UniversalBank.csv")
dim(x)





t(t(names(x))) #transpose of the dataframe


```


## droping the "id" and "zip" attributes for the dataset 


```{r}

new_x <-x[,-c(1,5)]
dim(new_x)

```

## converting education attribute from int to char 
```
```
```{r}
new_x$Education <- as.factor(new_x$Education)

```


## creating the dummy variables for the "education" attribute
```{r}
dumy <- dummyVars(~.,data=new_x)
the_neww <- as.data.frame(predict(dumy,new_x))
```

## Partitioning the data into training (60%) and validation (40%) set and setting the seed as we need to re-run the code.

```{r}
set.seed(1)
train.df <- sample(row.names(the_neww), 0.6*dim(the_neww)[1])
valid.df <- setdiff(row.names(the_neww),train.df)
t.df <- the_neww[train.df,]
v.df<- the_neww[valid.df,]
t(t(names(t.df)))

summary(t.df)

cat("The size of the training dataset is:",nrow(t.df))

```


```{r}

summary(v.df)

cat("The size of the validation dataset is:",nrow(v.df))
```



## normalizing the dataset
```{r}
train.norm <- t.df[,-10]
valid.norm <- v.df[,-10]

norm <- preProcess(t.df[,-10],method=c("center","scale"))

train.norm <- predict(norm,t.df[,-10])
valid.norm <- predict(norm,v.df[,-10])

```
## Questions

Consider the following customer:

##  1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?


## Creating new customer data
```{r}
new.cust <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer dataset
cust.norm <- predict(norm, new.cust)



```


## Performing kNN classification
```{r}
pred1 <- class::knn(train = train.norm, 
                       test = cust.norm, 
                       cl = t.df$Personal.Loan, k = 1)

pred1

```

## 2.What is a choice of k that balances between overfitting and ignoring the predictor information?
```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider
accu      <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  kn <- class::knn(train = train.norm, 
                   test = valid.norm, 
                   cl = t.df$Personal.Loan, k = i)
  accu[i, 2] <- confusionMatrix(kn, 
                  as.factor(v.df$Personal.Loan),positive = "1")$overall[1]
}

which(accu[,2] == max(accu[,2])) 


accu


```

```{r}
plot(accu$k,accu$overallaccuracy)
```
##### 3. Show the confusion matrix for the validation data that results from using the best k.


## confusion matrix
```{r}
predt <- class::knn(train = train.norm,
                   test = valid.norm,
                   cl =  t.df$Personal.Loan, k=3)

confusionMatrix(predt,as.factor(v.df$Personal.Loan))

```
##   4. Consider the following customer: Age = 40, Experience = 10, Income = 84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and CreditCard = 1. Classify the customer using the best k.


##  now creating the 2nd new customer  dataset
```{r}
customer2   <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1)


#Normalizing the 2nd customer dataset

cust_2 <- predict(norm , customer2)

```




##  Question-5: Repeating the process by partitioning the data into three parts - 50%, 30%, 20%,Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.

```{r}
set.seed(123)
Train_In <- sample(row.names(the_neww), .5*dim(the_neww)[1])#create train index

#create validation index
Va_In <- sample(setdiff(row.names(the_neww),Train_In),.3*dim(the_neww)[1])

Test_In =setdiff(row.names(the_neww),union(Train_In,Va_In))#create test index

train.d <- the_neww[Train_In,]

cat("The size of the  new training dataset is:", nrow(train.d))

```

```{r}
valid.d <- the_neww[Va_In, ]
cat("The size of the new validation dataset is:", nrow(valid.d))

```



```{r}
test.d <- the_neww[Test_In, ]
cat("The size of the new test dataset is:", nrow(test.d))
```


## Data Normalizing 

```{r}

norm.val <- preProcess(train.d[, -10], method=c("center", "scale"))
train.norm <- predict(norm.val, train.d[, -10])
valid.norm <- predict(norm.val, valid.d[, -10])
test.norm <- predict(norm.val, test.d[,-10])


```

##  Performing kNN and creating confusion matrix on training, testing, validation data


```{r}
pred31 <- class::knn(train = train.norm,
                    test = test.norm,
                    cl = train.d$Personal.Loan, k=3)

confusionMatrix(pred31,as.factor(test.d$Personal.Loan))
```



```{r}
pred41 <- class::knn(train = train.norm,
                    test = valid.norm,
                    cl = train.d$Personal.Loan, k=3)

confusionMatrix(pred41,as.factor(valid.d$Personal.Loan))


```


```{r}
pred51 <- class::knn(train = train.norm,
                       test = train.norm,
                       cl = train.d$Personal.Loan, k=3)

confusionMatrix(pred51,as.factor(train.d$Personal.Loan))


```


```